---
title: "(Very) basic steps on how to weight a survey sample"
subtitle: "An explained example with R code"
author: "JosepER"
output: html_notebook
---

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(forcats)
library(foreign)
library(magrittr)
library(tidyverse)
options(scipen = 9999)
```


## Introduction

### Note and bibliography 

This is a short introductory guide that shows to weight a survey. Please keep in mind that
there are many different ways of weighting a survey. This guide intends to be a practical document and
intentionally avoids explaining complex or advanced methods. Instead, I aim at providing 
a standard way of weighting and a limited number of variations.

For more information you can check the following introductory texts:

  * Valliant et al (2013) *Practical Tools for Designing and Weighting Survey Samples*. New York: Springer Science+Business Media.
  * Lohr (2009) *Sampling: Design and Analysis*. 2nd Edition. Boston: Books/Cole.
  
And the book accompaining the R 'survey' package:  

  * Lumley (2010) *Complex Surveys: A Guide to Analysis Using R*. New Jersey: John Wiley & Sons Inc.

It is important to note that this guide focuses on surveys based on 'probability sample'.
These are surveys where all units in our statistical population have a chance of being selected
and the probability of selection is known to the researcher.

### Basic steps in weighting a survey

Weighting is done to reduce survey bias. Briefly explained, weighting consists on making our sample of survey respondents (more) representative of our statistical population. By statistical population I mean all those units for which we want to compute estimates.

There are four basic steps in weighting. These are:

1. __Base/design weights__
2. __Non-response weights__
3. __Use of auxiliary data/calibration__
4. __Analysis of weight variability/trimming__

The first step consists on computing weights to take into account the differences of units in the probability of being sampled.
'Being sampled' means being selected from the survey frame (i.e. the list of all units) to be approached for a survey response.
This step can be skipped if all units in the survey frame have the same probability of being sampled. This happens, for example
1) when all units in the survey frame are approached for the sample or; 2) when the sampling design corresponds to either 'simple random sampling without replacement' or 'stratified random sampling without replacement' and the distribution of sampled units across stratums is proportional to the number of units in each stratum. These are called 'self-weighted' surveys. 

Second, we need to adjust our responses by the differences in probabilities of sampled units to reply to our survey. Our estimates would be biased if some profile of sampled units had higher propensity to reply than another and these profiles had differences in the dependent variables (i.e. our variables of interest). In this step, we need to estimate the probability of response using information available for both respondents and non-respondents. Non-response adjustment is not needed if all sampled units responded to the survey.

The third step consists on adjusting our weights using available information about total population estimates. Note that this requieres data that is different from that needed in non-response adjustment (second step). Here we need auxiliary data which tells us information (i.e. estimates such as proportions, means, sums, counts) about the statistical population. The same variables should be available from our respondents but here we don't need information about non-respondents.

The last step is to check the variablity in our computed weights. High variation in weights can lead to some observations having too much importance in our sample. Even if weights reduce bias, they might largely inflate variance of estimates. Therefore, some survey practitioners worry about dealing with highly unequal weights.


## Read data and data management

#### Import data

We first import data into R.

```{r echo=TRUE, warning=FALSE, message=FALSE}

sample.data <- read.spss("data/ESS7SDDFe1_1.sav", to.data.frame = T)  %>%
  filter(cntry == "United Kingdom")

paradata <- read.spss("data/ess7CFe02_1.sav", to.data.frame = T) %>%
  filter(cntry == "United Kingdom") 

responses <- read.spss("data/ESS7e02_1.sav", to.data.frame = T) %>%
  filter(cntry == "United Kingdom") 

```

#### Select variables

We select the variables we are going to use in our analysis. Here we just
write the names of the variables we intend to use.

```{r, echo=TRUE, warning=FALSE, message=FALSE}

vars.sample.data <- c("idno", "psu", "prob")

# Note: In sample file, domain and stratify variables could be useful for other countries.

vars.paradata <- c("idno", "typesamp", "interva", "telnum", 
                   "agea_1", "gendera1", "type", "access", 
                   "physa", "littera", "vandaa")

# Note: In paradata file, age of sampled unit could be useful for other countries.    

resp.id <- c("idno")

resp.y <- c("cgtsmke", "cgtsday",
         "alcfreq", "alcwkdy", "alcwknd")

resp.x <- c("vote", "prtvtbgb",
            "prtclbgb", "prtdgcl",
            "ctzcntr", "ctzshipc",
         "brncntr","cntbrthc",
         "gndr", "agea", "hhmmb","eisced",
         "pdwrk", "edctn", "uempla", "uempli", "rtrd",
         "wrkctra", "hinctnta")

```

Then we keep the variable labels (although these are not common in R).

```{r, echo=TRUE, warning=FALSE, message=FALSE}

selected.labels.sample.data <- attributes(sample.data)$variable.labels[which(names(sample.data) %in% vars.sample.data)]

selected.labels.paradata <- attributes(paradata)$variable.labels[which(names(paradata) %in% vars.paradata)]

selected.labels.responses <- attributes(responses)$variable.labels[which(names(responses) %in% c(resp.y, resp.x))] 

attributes(responses)$variable.labels %>% 
  cbind(names(responses),.) %>% 
  as_data_frame %>% 
  write_csv("interim_output/variable_labels.csv")

```

Now we do the selection of variables from the three data sets. 

```{r, echo=TRUE, warning=FALSE, message=FALSE}

sample.data %<>% 
  .[vars.sample.data]

paradata %<>%
  .[vars.paradata]

responses %<>%
  .[which(names(responses) %in% c(resp.id, resp.y, resp.x))]

```


### Merging datafiles

We merge the 'paradata' file containing all sampled units with the 'survey responses' file.
In a real situation, there would also be a 'survey frame' dataset. This would be a third data file
from which the sample was selected. Ideally, the 'survey frame' would include all units from the 
population.

```{r, echo=TRUE, warning=FALSE, message=FALSE}

data <- paradata %>%
  left_join(sample.data, by = "idno") %>%
  left_join(responses, by = "idno") %>%
  arrange(interva)

rm(paradata,
   sample.data,
   responses)

```

And we add the labels we kept before.

```{r, echo=TRUE, warning=FALSE, message=FALSE}

# Note: There are other variables that could be selected from paradata (i.e. result of visits, refusals, etc.)

attributes(data)$variable.labels <- c(selected.labels.paradata, selected.labels.sample.data[!names(selected.labels.sample.data) %in% "idno"],
                                      selected.labels.responses)
```


### Recoding

Recoding daily cigarette consumption. All those that responded that don't smoke should have a 0.
```{r, echo=TRUE, warning=FALSE, message=FALSE}
data$cgtsday[data$cgtsmke %in% c("I have never smoked",
                                 "I don't smoke now but I used to",
                                 "I have only smoked a few times")] <- 0

data$alcohol_day <- NA 
data$alcohol_day <- (data$alcwkdy * 5 + data$alcwknd *2)/7 

data$alcohol_day[which(data$alcfreq == "Several times a week")] <- data$alcohol_day / 2.5
data$alcohol_day[which(data$alcfreq == "Once a week")] <- data$alcohol_day/7
data$alcohol_day[which(data$alcfreq == "2-3 times a month")] <- data$alcohol_day/10
data$alcohol_day[which(data$alcfreq == "Once a month")] <- data$alcohol_day/30
data$alcohol_day[which(data$alcfreq == "Less than once a month")] <- data$alcohol_day/50
data$alcohol_day[which(data$alcfreq == "Never")] <- 0

resp.y <- c(resp.y, "alcohol_day")
```


## Exploring and presenting the dataset

The merged data set contains sampled **respondents and non-respondents**. It contains a total of `r dim(data)[[1]]` units and `r dim(data)[[2]]` variables.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
dim(data)
```

The data set contains information about `r nrow(data %>% filter(interva == "Complete and valid interview related to CF"))` respondents and `r nrow(data %>% filter(interva != "Complete and valid interview related to CF"))` non-respondents.


And this is a list of the variables it contains (with their labels). **idno** is the individual identification variable.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
cbind(names(data),attributes(data)$variable.labels) %>% 
  as_data_frame() %>%
  print()
```

Here I will try to give population estimates for cigarette and alcohol consumption in the UK. These will be our 'y' variables or variables of interest. The information for these variables was obtained from survey responses. The idea is to first give descriptives of the distribution of these two variables such as quantiles and mean. Then I will do a simple extrapolation and compute total cigarette and alcohol consumption for the whole UK.

These are our y variables:

* __cgtsday__: Number of cigarettes smoked on a typical day.
* __alcohol_day__: Grams of alcohol ingested on a daily basis. Computed in the *Recoding* section from the amount of alcohol drank last time during weekdays and weekend.


```{r, echo=TRUE, warning=FALSE, message=FALSE}
data[resp.y]
```


#### Paradata variables

Our survey contains variables which give information about the data collection process. First, we have some variables that come from the 'sample data (SDDF)' file. These contain info about the 'primary sampling unit' and the probability of each unit of being selected in the sample. These two variables are only available for respondents. In most projects we would have to compute the probability of being sampled by ourselves.   

* __psu__: This variable includes information on the primary sampling unit (PSU). In the UK this refered to the 'postcode address file'.
* __prob__: Probability of being included in the sample (i.e. approached for survey).

```{r, echo=TRUE, warning=FALSE, message=FALSE}
data[vars.sample.data]
```

The dataset also contains variables for all sampled units (i.e. respondents and non-respondents). These give information  about the events that occurred during the data collection process. We will use these variables as covariates during the computation of **Non-response weights** in step two.  

* __typesamp__: Refers to the type of unit sampled. In the UK, all units sampled were addresses. In other countries, households and individual people were sampled.   
* __interva__: Shows the final outcome of the contact. In the UK sample, only codes 'Complete ...' and 'No interview ...' were used.
* __telnum__: The interviewed person gave his/her mobile phone to the interviewer.
* __agea_1__:  Interviewer estimation of age of respondent or household member who refuses to give the interview.
* __gendera1__: Interviewer estimation of gender of respondent or household member who refuses to give the interview.
* __type__: Type of house sampled unit lives in. 
* __access__: Entry phone or locked gate/door before reaching respondent's individual door.
* __physa__: Interviewer assessment overall physical condition building/house.
* __littera__: Interviewer assessment of amount of litter and rubbish in the immediate vicinity.
* __vandaa__: Interviewer assessment of amount of vandalism and graffiti in the immediate vicinity.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
data[vars.paradata] 
```


#### Survey responses

Apart from the variables of interest (cigarette and alcohol consumption) our dataset has other variables obtained from survey responses. Obviously, these are only available for respondents. We will try to use these variables to calibrate the survey in __Use of auxiliary data/calibration__ step. Some of these variables are:

* __vote__: Voted last national election (Yes/No)
* __prtvtbgb__: Party voted for in last national election
* __prtclbgb__: Which party feel closer to, United Kingdom
* __prtdgcl__: How close does the repondent feel to the party party from 'prtclbgb'
* __ctzcntr__: Has UK citizenship (Yes/No)
* __ctzshipc__: Citizenship of respondent
* __brncntr__:Respondent born in the UK
* __cntbrthc__: Respondent country of birth
* __gndr__: Gender of respondent
* __agea__: Calculated age of respondent
* __eisced__: Highest level of education of respondent
* __pdwrk__: In paid work
* __edctn__: In education
* __uempla__: In unemployment, actively looking for a job
* __uempli__: In unemployment, not actively looking for a job
* __rtrd__: Retired
* __wrkctra__: Employment contract unlimited or limited duration
* __hinctnta__: Household's total net income, all sources


```{r, echo=TRUE, warning=FALSE, message=FALSE}
data[c(resp.id, resp.x)] 
```


This is how the whole dataset looks like:


```{r, echo=TRUE, warning=FALSE, message=FALSE}
head(data, 10)
```

### Descriptives

(It would be a good practice to give descriptives for the main variables.)

##### Variables of interest

##### Covariates

##### Auxiliary data


## Step 1: Design weights

The first step in weighing is to take into account the differences in the probability of being selected for the sample. The European Social Survey (ESS) did not use a register of people in the UK (in other countries they did). They first selected postcode sectors from the Post Office’s small user
postcode address file (PAF), merging smaller sectors. The probability of each PAF of being selected was proportional to the number of addresses it contained. Then, they selected 20 addresses inside of each sampled PAF and a dwelling for each address. For each dewlling they selected a household and then a person in each household. The full explanation of the sampling procedure is given in page 163 of The data documentation report ([Edition 3.1](http://www.europeansocialsurvey.org/docs/round7/survey/ESS7_data_documentation_report_e03_1.pdf) ). 

This sampling design is most probably done because they had a list of addresses but not households or individuals. If we don't weight this survey, we would probably over-represent people in addresses that have smaller number of dewellings, dwellings that include smaller number of households and households that comprise smaller number of people.

Fortunately for us, the probablity of each respondent of being sampled was included in the ESS dataset. In many projects, however, we will have to compute sampling probailities ourselves. A basic but **important test** that should be performed after computing the probabilities is making sure that all probabilities are between 0 and 1.

```{r, echo=TRUE, warning=FALSE, message=FALSE}

probabilities <- data %>%
  summarise(min.probability = min(prob, na.rm = T),
            max.probability = max(prob, na.rm = T)) %>%
  as.vector()

print(probabilities)

if(probabilities$min.probability < 0){stop("Minimum probability of being sampled is smaller than 0. Review sampling probabilities before computing base weights.")}else if(probabilities$max.probability > 1){stop("Maximum probability of being sampled is larger than 1. Review sampling probabilities before computing base weights.")}

rm(probabilities)

```


We see that there are actually `r length(unique(data$prob))` unique sampling probabilities computed in the dataset. 

```{r, echo=TRUE, warning=FALSE, message=FALSE}
unique(sort(data$prob))

```

The vast majority of respondents had probability of `r sort(table(data$prob), decreasing = T)[[1]]` or `r sort(table(data$prob), decreasing = T)[[2]]`. This would mean that the most frequent probabilites of inclusion were of `r 1/round(sort(table(data$prob), decreasing = T)[[1]],0)` and `r 1/round(sort(table(data$prob), decreasing = T)[[2]],0)` respectively.
We see a minority of around 15% of observations with smaller probabilities. 

```{r, echo=TRUE, warning=FALSE, message=FALSE}
ggplot(data, aes(x = prob)) +
  geom_histogram()

```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
table(round(data$prob*100,6))
```


We see that 


```{r, echo=TRUE, warning=FALSE, message=FALSE}

data %>%
  group_by(type) %>%
  summarise(n = n(),
    mean.prob.percentage = mean(prob, na.rm = T)*100) %>%
      arrange(desc(mean.prob.percentage))

```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
data %<>%
  mutate(hhmmb.factor = as.factor(hhmmb) %>% fct_recode(`+5` = "6",
                                                        `+5` = "7",
                                                        `+5` = "8"))

data %>%
  filter(!is.na(hhmmb.factor)) %>%
  group_by(hhmmb.factor) %>%
  summarise(n = n(),
    mean.prob.percentage = mean(prob, na.rm = T)*100) %>%
      arrange(desc(mean.prob.percentage))

```

The base weights are equal to the inverse of the probability of inclusion. Therefore, the base weight (*d_{0}*) of a respondent (*i*) will be equal to: $d_{0i} =  1/\pi_{i}$. 

```{r, echo=TRUE, warning=FALSE, message=FALSE}

data %<>%
  mutate(base.weight = 1/prob)

data %>%
  select(prob, base.weight)


```

Here we compute the base weight from the probability given in the ESS database.   

```{r, echo=TRUE, warning=FALSE, message=FALSE}

data %<>%
  mutate(base.weight = 1/prob)

data %>%
  select(prob, base.weight)

```

A simple interpretation of base weights it 'the number of units in our population that each unit in our sample represents'. There is a simple but **important test** that we should perform after computing base weights. The sum of all base weights should be equal to the total number of units in our population. The ESS dataset for UK only included sampling probabilities for respondents (i.e. sampled units that responded to the survey!) but they did not include sampling probabilities of non-respondents.  I would guess that this is because sampling probability depends on information that is obtained from the interview (i.e. number of people in household, number of households in dwelling, etc.).  Not knowing the sampling probability for some sampled units is not an optimal situation. 

The sum of our computed weights in the ESS dataset with `r table(data$interva)[["Complete and valid interview related to CF"]]` respondents equals `r data %>%  summarise(sum.base.weights.ess.dataset = round(sum(base.weight, na.rm = T),0)) %>% .[["sum.base.weights.ess.dataset"]]`. Doing a very simple Extrapolation to include the `r table(data$interva)[["No interview for other reason"]]` non-respondents would give us a sum of weights equal to `r data %>%  summarise(sum.base.weights.ess.dataset = round(sum(base.weight, na.rm = T),0)) %>% .[["sum.base.weights.ess.dataset"]]* (nrow(data)/ table(data$interva)[["Complete and valid interview related to CF"]])`. This last figure would be much closer to the total UK population over 15. 



```{r, echo=TRUE, warning=FALSE, message=FALSE}

data %>%
  summarise(sum.base.weights.ess.dataset = sum(base.weight, na.rm = T))

```


## Step 2: Non-response weights

## Step 3: Use of auxiliary data

## Step 4: Analysis of weight variability

## Computing an estimate


